### 2017-12-04 Presentation/MSDS6306_CaseStudy_2.Rmd
RMD First Draft.

---
title: "MSDS6306_CaseStudy_2"
author: "Megan Hodges and Vivek Viswanathan"
date: "December 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages required
```{r}
if (!require(tidyverse)) install.packages("tidyverse")
if (!require(repmis)) install.packages("repmis")
if (!require(dplyr)) install.packages("dplyr")
if (!require(tidyr)) install.packages("tidyr")
if (!require(ggplot2)) install.packages("ggplot2")
if (!require(sqldf)) install.packages("sqldf")
if (!require(kimisc)) install.packages("kimisc")
if (!require(XML)) install.packages("XML")
if (!require(RCurl)) install.packages("RCurl")
if (!require(rvest)) install.packages("rvest")
if (!require(plyr)) install.packages("plyr")
if (!require(pastecs)) install.packages("pastecs")
```

# Load Libraries
```{r}
library(repmis)
library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(sqldf)
library(kimisc)
library(XML)
library(RCurl)
library(rvest)
library(plyr)
library(pastecs)
```

# R - environment
```{r}
sessionInfo()
```

# Set Base Directory:

```{r}
# Executing from Vivek's System:
BaseDir <- "C:/Vivek/Data_Science/MSDS6306-DoingDataScience/Case_Study_2/MSDS6306_CaseStudy_2/"

# Executing from Megan's System:
# BaseDir <- "/Users/megandiane/Desktop/DDS_Class/Case_Study_2/MSDS6306_CaseStudy_2"
```

# Set Other Working Directories and File Path
```{r}
DataDir <- paste(BaseDir,"Data", sep = "/")
CodeDir <- paste(BaseDir,"Code", sep = "/")
PresenatationDir <- paste(BaseDir,"Presentation", sep = "/")

ProcrastinationDataFile <- paste(DataDir, "Procrastination.csv", sep = "/")
```

# Read the Data File
```{r}
ProcrastinationData <- read.csv(ProcrastinationDataFile, sep = ",", header = T, na.strings = "")

dim(ProcrastinationData)
```

# Assign Column Names
```{r}
names(ProcrastinationData) <- c("Age", "Gender", "Kids", "Education", "WorkStatus", 
                                "AnnualIncome", "CurrOccption", "PostHeldYrs", "PostHeldMths", 
                                "CmmuntySize", "CntryResdnc", "MaritlStatus", "SonsCnt", 
                                "DaughtersCnt", "D1DsnTmeWst", "D2DelayActn", "D3HesitatDsn", 
                                "D4DelayDsn", "D5PutoffDsn", "A1BillsOnTm", "A2OnTm4Appt", 
                                "A3Rdy4NxtDy", "A4RuningLate", "A5ActvtDlyd", "A6TmeMgmtTrg", 
                                "A7FrndsOpn", "A8ImpTskOnTm", "A9MissDedlns", "A10RnOutOfTm", 
                                "A11DrAptOnTm", "A12MorPnctul", "A13RtneMntnc", "A14SchdulLte", 
                                "A15DldActCst", "G1LteToTsk", "G2LteTktPrch", "G3PlnPrtyAhd", 
                                "G4GetUpOnTme", "G5PstLtrOnTm", "G6RtrnCalls", "G7DlyEsyTsks", 
                                "G8PrmptDscsn", "G9DlyTskStrt", "G10TrvlRsh", "G11RdyOnTme", 
                                "G12StayOnTsk", "G13SmlBlOnTm", "G14PrmptRSVP", "G15TskCmpErl", 
                                "G16LstMntGft", "G17DlyEsntPr", "G18DyTskCmpl", "G19PshTskTmr", 
                                "G20CmpTskRlx", "S1LfClsI2dl", "S2LfCndExlnt", "S3StsfdWtLf", 
                                "S4GtImThgsLf", "S5LvAgChgNth", "CnsdrSlfProc", "OthCsndrProc")

```

# To Clean up the data from Procrastination.csv we took the "Male" and "Female" that we found under the Column of "Sons" and
# replaced it with the correct response of 1 = Male and 2 = Female.
```{r}
levels(ProcrastinationData$SonsCnt) <- c(levels(ProcrastinationData$SonsCnt), "1", "2")
ProcrastinationData$SonsCnt[ProcrastinationData$SonsCnt=='Male'] <- '1'
ProcrastinationData$SonsCnt[ProcrastinationData$SonsCnt=='Female'] <- '2'

ProcrastinationData[,"SonsCnt"] <- as.integer(as.character(ProcrastinationData[,"SonsCnt"]))
ProcrastinationData[,"PostHeldYrs"] <- as.integer(ProcrastinationData[,"PostHeldYrs"])
```

#To Clean up the data from Procrastination.csv we did the folowing:
#1. If Annual Income was blank, we filled with a -0.01, that way we could easily differentiate from true data, and keep as a numerical data type 
#2. If the Country of Residence was filled with a "0", we filled with "Missing", to keep as a character data type
#3. If Current Occupation was filled with "0" or "please specify", we filled with "Missing", to keep as a character data type

#4. For all of the Procrastination Scales: DP, AIP, GP, and SWLS, the means were taken and columns added to reflect those means.
#   This Mean data is to be used later on.
```{r}
ProcTrans <- ProcrastinationData %>% 
  mutate(AnnualIncome=replace(AnnualIncome, is.na(AnnualIncome), -0.01)) %>%
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  mutate_if(is.character, funs(ifelse(is.na(.), "Missing", .))) %>% 
  mutate_if(is.factor, funs(ifelse(is.na(.), "Missing", as.character(.)))) %>%
  mutate(CntryResdnc=replace(CntryResdnc, CntryResdnc=="0", "Missing")) %>%
  mutate(PostHeldYrs=replace(PostHeldYrs, PostHeldYrs==999, NA)) %>%
  mutate(CurrOccption=replace(CurrOccption, 
                              (CurrOccption=="0" | CurrOccption=="please specify"), 
                              "Missing")) %>%
  mutate(DPMean=rowMeans(ProcrastinationData[c('D1DsnTmeWst', 'D2DelayActn', 'D3HesitatDsn', 
                                               'D4DelayDsn', 'D5PutoffDsn')], na.rm=TRUE)) %>%
  mutate(AIPMean=rowMeans(ProcrastinationData[c('A1BillsOnTm', 'A2OnTm4Appt', 'A3Rdy4NxtDy', 
                                                'A4RuningLate', 'A5ActvtDlyd', 'A6TmeMgmtTrg', 
                                                'A7FrndsOpn', 'A8ImpTskOnTm', 'A9MissDedlns', 
                                                'A10RnOutOfTm', 'A11DrAptOnTm', 'A12MorPnctul', 
                                                'A13RtneMntnc', 'A14SchdulLte', 'A15DldActCst')], 
                          na.rm=TRUE)) %>%
  mutate(GPMean=rowMeans(ProcrastinationData[c('G1LteToTsk', 'G2LteTktPrch', 'G3PlnPrtyAhd', 
                                               'G4GetUpOnTme', 'G5PstLtrOnTm', 'G6RtrnCalls', 
                                               'G7DlyEsyTsks', 'G8PrmptDscsn', 'G9DlyTskStrt', 
                                               'G10TrvlRsh', 'G11RdyOnTme', 'G12StayOnTsk', 
                                               'G13SmlBlOnTm', 'G14PrmptRSVP', 'G15TskCmpErl', 
                                               'G16LstMntGft', 'G17DlyEsntPr', 'G18DyTskCmpl', 
                                               'G19PshTskTmr', 'G20CmpTskRlx')], na.rm=TRUE)) %>% 
  mutate(SWLSMean=rowMeans(ProcrastinationData[c('S1LfClsI2dl', 'S2LfCndExlnt', 'S3StsfdWtLf', 
                                                 'S4GtImThgsLf', 'S5LvAgChgNth')], na.rm=TRUE))

ProcTrans[grep("(teacher)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Teacher"
ProcTrans[grep("(accountant)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Accountant"
ProcTrans[grep("(professor)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Professor"
ProcTrans[grep("(engineer)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Engineer"
ProcTrans[grep("(manager)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Manager"
ProcTrans[grep("(retired)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Retired"
ProcTrans[grep("(marketing)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Marketing"
ProcTrans[grep("(attorney)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Attorney"
ProcTrans[grep("(doctor)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Doctor"
ProcTrans[grep("(student)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Student"
ProcTrans[grep("(writer)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Writer"
ProcTrans[grep("(developer)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Developer"
ProcTrans[grep("(nurse)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Nurse"
ProcTrans[grep("(consultant)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Consultant"
ProcTrans[grep("(scientist)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Scientist"
ProcTrans[grep("(admin)+", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Administrator"
ProcTrans[grep("(Financial)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Finance"
ProcTrans[grep("(director)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Director"
ProcTrans[grep("(customer service)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Customer Service"
ProcTrans[grep("(libra)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Librarian"
ProcTrans[grep("(Business / Test Analyst)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "BA"
ProcTrans[grep("(Business Systems Analyst)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "BA"
ProcTrans[grep("(BUSINESS CONSULTA)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "BA"
ProcTrans[grep("(busi)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Business Owner"
ProcTrans[grep("(artist)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Artist"
ProcTrans[grep("(web)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Website Designer"
ProcTrans[grep("(prog)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Programmer"
ProcTrans[grep("(vp)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Vice President"
ProcTrans[grep("(train)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Trainer"
ProcTrans[grep("(technician)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Technician"
ProcTrans[grep("(CTO))", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "CTO"
ProcTrans[grep("(technici)", ProcTrans$CurrOccption, ignore.case = T), "CurrOccption"] <- "Technician"

sapply(ProcTrans, class)
```

#Scrap Data from Wikipedia: List of Countries By Human Development Index
```{r}
HumDevUrl <- "https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index#Complete_list_of_countries"
```

#Very High Human Development
```{r}
VHighHumDev <- HumDevUrl %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[5]/table') %>%
  html_table(fill = T)

VHighHumDev <- data.frame(VHighHumDev[[1]])
```

#Remove Unnecessary Rows
```{r}
VHighHumDev_1Head <- VHighHumDev[-c(1,2,3,30,31), ]

#Remove Unnecessary Columns
```{r}
VHighHumDev[2] <- list(NULL)
VHighHumDev_Clean <- VHighHumDev_1Head[,c(1,3,4)]
```

#Rename Columns
```{r}
VHighHumDev <- rename(VHighHumDev_Clean, c("X1"="Rank", "X3"="Country", "X4"="HDI"))

VHighHumDev$HumDev_Categ <- "VHigh"

VHighHumDev[,"HDI"] <- as.numeric(VHighHumDev[,"HDI"])
```

# Remove Unused Environment Variables
```{r}
rm(VHighHumDev_1Head, VHighHumDev_Clean)
```

#High Human Development
```{r}
HighHumDev <- HumDevUrl %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[6]/table') %>%
  html_table(fill = T)

HighHumDev <- data.frame(HighHumDev[[1]])
```

#Remove Unnecessary Rows
```{r}
HighHumDev_1Head <- HighHumDev[-c(1,2,3,32,33), ]
```

#Remove Unnecessary Columns
```{r}
HighHumDev[2] <- list(NULL)
HighHumDev_Clean <- HighHumDev_1Head[,c(1,3,4)]
```

#Rename Columns
```{r}
HighHumDev <- rename(HighHumDev_Clean, c("X1"="Rank", "X3"="Country", "X4"="HDI"))

HighHumDev$HumDev_Categ <- "High"

HighHumDev[,"HDI"] <- as.numeric(HighHumDev[,"HDI"])
```

# Remove Unused Environment Variables
```{r}
rm(HighHumDev_1Head, HighHumDev_Clean)
```

#Medium Human Development
```{r}
MedHumDev <- HumDevUrl %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[7]/table') %>%
  html_table(fill = T)

MedHumDev <- data.frame(MedHumDev[[1]])
```

#Remove Unnecessary Rows
```{r}
MedHumDev_1Head <- MedHumDev[-c(1,2,3,24,25), ]
```

#Remove Unnecessary Columns
```{r}
MedHumDev[2] <- list(NULL)
MedHumDev_Clean <- MedHumDev_1Head[,c(1,3,4)]
```

#Rename Columns
```{r}
MedHumDev <- rename(MedHumDev_Clean, c("X1"="Rank", "X3"="Country", "X4"="HDI"))

MedHumDev$HumDev_Categ <- "Med"

MedHumDev[,"HDI"] <- as.numeric(MedHumDev[,"HDI"])
```

# Remove Unused Environment Variables
```{r}
rm(MedHumDev_1Head, MedHumDev_Clean)
```

#Low Human Development
```{r}
LowHumDev <- HumDevUrl %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div/div[8]/table') %>%
  html_table(fill = T)

LowHumDev <- data.frame(LowHumDev[[1]])
```

#Remove Unnecessary Rows
```{r}
LowHumDev_1Head <- LowHumDev[-c(1,2,3,25,26), ]
```

#Remove Unnecessary Columns
```{r}
LowHumDev[2] <- list(NULL)
LowHumDev_Clean <- LowHumDev_1Head[,c(1,3,4)]
```

#Rename Columns
```{r}
LowHumDev <- rename(LowHumDev_Clean, c("X1"="Rank", "X3"="Country", "X4"="HDI"))

LowHumDev$HumDev_Categ <- "Low"

LowHumDev[,"HDI"] <- as.numeric(LowHumDev[,"HDI"])
```

# Remove Unused Environment Variables
```{r}
rm(LowHumDev_1Head, LowHumDev_Clean)
```

#Combine the Four Data Frames
```{r}
Total_HumDev <- rbind(VHighHumDev, HighHumDev, MedHumDev, LowHumDev)

Total_HumDev <- within(Total_HumDev, rm("Rank"))
```

# Remove Unused Environment Variables
```{r}
rm(VHighHumDev, HighHumDev, MedHumDev, LowHumDev)
```

#Merge this data frame to the Country of Residence column of Procrastination.csv so that your data now has an HDI column and HDI categories
```{r}
Merged_ProctransHumDev <- merge(ProcTrans, Total_HumDev, by.x=c("CntryResdnc"), by.y = c("Country"))
```

#Remove all Participants who are under 18
#Also chose to remove all Age of Zero (0) because our client is looking for Procrastination as it relates to positions held, how long, and annual income, and all observations with Zero Age, also did not have jobs listed
```{r}
Merged_ProctransHumDev <- Merged_ProctransHumDev[Merged_ProctransHumDev$Age >= 18.0, ]

unique(format(Merged_ProctransHumDev$Age, digits = 10))

Merged_ProctransHumDev_DescripStats <- Merged_ProctransHumDev[c("Age", "AnnualIncome", "HDI",
                                                                "DPMean", "AIPMean",
                                                                "GPMean", "SWLSMean")]

options(scipen=100)
options(digits=2)
ProctransHumDev_DescripStats <- stat.desc(Merged_ProctransHumDev_DescripStats)
ProctransHumDev_DescripStats

hist(Merged_ProctransHumDev_DescripStats$DPMean, 
     xlab = "DPMean", ylab = "Occurence", 
     main = "Distribution of Occurence of DPMean")
hist(Merged_ProctransHumDev_DescripStats$GPMean, 
     xlab = "DPMean", ylab = "Occurence", 
     main = "Distribution of Occurence of GPMean")
```

# Remove Unused Environment Variables
```{r}
rm(Merged_ProctransHumDev_DescripStats, ProctransHumDev_DescripStats)

Cnt_By_Gender <- as.data.frame(table(Merged_ProctransHumDev$Gender))
Cnt_By_Gender
Cnt_By_WorkStatus <- as.data.frame(table(Merged_ProctransHumDev$WorkStatus))
Cnt_By_WorkStatus
Cnt_By_Curr_Occupation <- as.data.frame(table(Merged_ProctransHumDev$CurrOccption))
Cnt_By_Curr_Occupation %>%
  arrange(desc(Freq)) %>%
  head(30)

Cnt_By_CntryResdnc <- Merged_ProctransHumDev %>%
  group_by(CntryResdnc) %>% 
  do(data.frame(nrow=nrow(.))) %>%
  arrange(desc(nrow))

Cnt_By_CntryResdnc

Merged_ProctransHumDev %>%
  group_by(CnsdrSlfProc, OthCsndrProc) %>% 
  do(data.frame(nrow=nrow(.))) %>%
  arrange(desc(nrow)) %>%
  filter((CnsdrSlfProc == "yes" & OthCsndrProc == "yes") 
         | (CnsdrSlfProc == "no" & OthCsndrProc == "no"))

DP_Top15 <- aggregate(DPMean ~ CntryResdnc+HDI, Merged_ProctransHumDev, mean) %>%
  arrange(desc(DPMean)) %>%
  head(n=15)

ggplot(data = merge(Merged_ProctransHumDev, within(DP_Top15, rm("DPMean", "HDI")), by = "CntryResdnc")) +
  geom_bar(aes(x=reorder(CntryResdnc,DPMean,mean), DPMean, fill = HumDev_Categ),
           stat = "summary", fun.y = "mean", show.legend = T) + 
  xlab("Country") + ylab("DPMean") + 
  ggtitle("Top 15 Countries of DP Procrastination Mean Scale") + 
  theme(axis.text.x = element_text(hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip()

GP_Top15 <- aggregate(GPMean ~ CntryResdnc+HDI, Merged_ProctransHumDev, mean) %>%
  arrange(desc(GPMean)) %>%
  head(n=15)

ggplot(data = merge(Merged_ProctransHumDev, within(GP_Top15, rm("GPMean", "HDI")), by = "CntryResdnc")) +
  geom_bar(aes(x=reorder(CntryResdnc,GPMean,mean), GPMean, fill = HumDev_Categ),
           stat = "summary", fun.y = "mean", show.legend = T) + 
  xlab("Country") + ylab("GPMean") + 
  ggtitle("Top 15 Countries of GP Procrastination Mean Scale") + 
  theme(axis.text.x = element_text(hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) + ylim(0,5) +
  coord_flip()

GP_DP_Common_Cntry <- as.data.frame(intersect(DP_Top15$CntryResdnc, GP_Top15$CntryResdnc))
dim(as.data.frame(GP_DP_Common_Cntry))
names(GP_DP_Common_Cntry) <- c("CommonCntry")
GP_DP_Common_Cntry

plot(Merged_ProctransHumDev$Age, Merged_ProctransHumDev$AnnualIncome, 
     xlab="Age", ylab="Annual Income", 
     main="Age vs Annual Income", pch=2, cex.main=1.5, 
     frame.plot=FALSE, col=ifelse(Merged_ProctransHumDev$Gender=="Male", "red", "blue"))

legend("topleft", pch=c(2,2), col=c("red", "blue"), 
       c("Male", "Female"), bty="o",  box.col="darkgreen", cex=.8)

ggplot(Merged_ProctransHumDev, aes(x=Age,y=AnnualIncome,color=Gender)) + 
  geom_point()+ geom_smooth(method = lm) + 
  xlab("Age") + ylab("Annual Income") + 
  ggtitle("Age vs Annual Income plot for all countries") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data = merge(Merged_ProctransHumDev, GP_DP_Common_Cntry, x.by = "CntryResdnc", y.by = "CommonCntry")) +
  aes(x=Age,y=AnnualIncome,color=Gender) + 
  geom_point()+ geom_smooth(method = lm) + 
  xlab("Age") + ylab("Annual Income") + 
  ggtitle("Age vs Annual Income plot for the 8 countries \ncommon between Top 15 DP and GP countries") +
  theme(plot.title = element_text(hjust = 0.5))

plot(Merged_ProctransHumDev$SWLSMean, Merged_ProctransHumDev$HDI, 
     xlab="SWLSMean", ylab="HDI", ylim = c(0,1),
     main="SWLSMean vs HDI", pch=2, cex.main=1.5, 
     frame.plot=FALSE, col=ifelse(Merged_ProctransHumDev$Gender=="Male", "red", "blue"))

ggplot(Merged_ProctransHumDev, aes(x=SWLSMean,y=HDI,color=Gender)) + 
  geom_point()+ geom_smooth(method = lm) + 
  xlab("HDI Category") + ylab("SWLSMean") + 
  ggtitle("Scatterplot of Mean Life Satisfaction by \nHuman Development Index Category by Gender") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(Merged_ProctransHumDev) +
  geom_bar(aes(x=reorder(HumDev_Categ,-SWLSMean,mean), SWLSMean, fill = HumDev_Categ),
           stat = "summary", fun.y = "mean", show.legend = T) + 
  xlab("HDI Category") + ylab("SWLSMean") + 
  ggtitle("Bar Chart of Mean Life Satisfaction by \nHuman Development Index Category") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

LS_Top15 <- aggregate(SWLSMean ~ CntryResdnc+HDI, Merged_ProctransHumDev, mean) %>%
  arrange(desc(SWLSMean)) %>%
  head(n=15)

ggplot(data = merge(Merged_ProctransHumDev, within(LS_Top15, rm("SWLSMean", "HDI")), by = "CntryResdnc")) +
  geom_bar(aes(x=reorder(CntryResdnc,SWLSMean,mean), SWLSMean, fill = HumDev_Categ),
           stat = "summary", fun.y = "mean", show.legend = T) + 
  xlab("Country") + ylab("SWLSMean") + 
  ggtitle("Top 15 Countries of LS Procrastination Mean Scale") + 
  theme(axis.text.x = element_text(hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) + ylim(0,5) +
  coord_flip()
```

#Write the Human Development Data to CSV
```{r}
HumDevDataFile <- paste(DataDir, "HumanDevelopment.csv", sep = "/")
write.table(Total_HumDev,HumDevDataFile,row.names=F, col.names = T, sep = ",")
```

#Write the Cleaned Input Data with HDI to CSV
```{r}
CleanInputData_w_HDI_File <- paste(DataDir, "CleanedInput_w_HDI.csv", sep = "/")
write.table(Merged_ProctransHumDev,CleanInputData_w_HDI_File,row.names=F, col.names = T, sep = ",")
```

#Write the Human Development Data to CSV
```{r}
Top_15_DP <- paste(DataDir, "Top_15_DP_Cntry.csv", sep = "/")
write.table(DP_Top15, Top_15_DP,row.names=F, col.names = T, sep = ",")
```

#Write the Cleaned Input Data with HDI to CSV
```{r}
Top_15_GP <- paste(DataDir, "Top_15_GP_Cntry.csv", sep = "/")
write.table(GP_Top15, Top_15_GP,row.names=F, col.names = T, sep = ",")
```

# Remove Unused Environment Variables
```{r}
rm(Total_HumDev, Merged_ProctransHumDev, DP_Top15, GP_Top15, LS_Top15, GP_DP_Common_Cntry, 
   Cnt_By_CntryResdnc, Cnt_By_Curr_Occupation, Cnt_By_Gender, Cnt_By_WorkStatus, ProcTrans,
   ProcrastinationData, CleanInputData_w_HDI_File, HumDevDataFile, Top_15_DP, Top_15_GP)

rm(BaseDir, CodeDir, DataDir, HumDevUrl, PresenatationDir, ProcrastinationDataFile)
```